---
title: "Lab 2 - R"
author: "Tomás León"
date: "1/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(RColorBrewer)
library(ggplot2)
library(ggmap)
library(dplyr)
```

First, we have to load our data. We have spatial state data and COVID data in table form.

We need to subset/modify to remove the cruiseships and to look at only the Continental US for mapping purposes.


```{r Read in Data, include = FALSE}
us_states <- st_read("cb_2019_us_state_500k//")
covid_states <- read.csv("covid_state_01-23-2021.csv")

us_states$NAME
covid_states$Province_State

## Remove cruise ships
covid_states <- covid_states[-c(10,14),]

## Create continental-only versions
non_continental_names <- c("Guam", "United States Virgin Islands", "Virgin Islands", "American Samoa", "Commonwealth of the Northern Mariana Islands", "Northern Mariana Islands", "Hawaii", "Alaska", "Puerto Rico")

covid_continental <- covid_states %>% filter(!(Province_State %in% non_continental_names))
us_continental <- us_states %>% filter(!(NAME %in% non_continental_names))

us <- c(left = -125, bottom = 25, right = -67, top = 49)
us_basemap <- get_stamenmap(us, zoom = 5, maptype = "toner-lite")
#us_basemap <- readRDS("us_basemap.Rds")
```

Now, let's look at what data we have.

```{r Examine Data}
colnames(us_continental)
colnames(covid_continental)
```

Let's try out some basic map plotting.

```{r Basic Map Plot}
plot(us_continental["NAME"], main = "Continental US")

ggmap(us_basemap)

ggmap(us_basemap) + geom_point(data = covid_continental, aes(x = Long_, y = Lat, size = Deaths),
                               col = "darkred") + theme_void()

ggmap(us_basemap) + geom_sf(data = us_continental, aes(fill = NAME), inherit.aes = FALSE) + guides(fill = FALSE)


```

Now let's join our two data sources so we can plot the information together.

```{r Do table join}

us_continental_covid <- left_join(us_continental, covid_continental, by = c(NAME = "Province_State"))

#us_continental_covid
```

After doing the join, we can now create choropleth maps for the US state COVID data.

```{r Make choropleth map with COVID state data}

ggplot(us_continental_covid) + geom_sf(aes(fill = Incident_Rate)) + 
  scale_fill_distiller(name = "Case Incidence", palette = "RdBu") + theme_void()


pal <-brewer.pal(5, "Greens")

plot(us_continental_covid["Testing_Rate"],
     main = "Testing Rate in US States",
     breaks = "pretty", nbreaks = 5, pal = pal)


```

